Resonant Scattering in a Uniform Sphere with Large Optical Depth

### 2021-07-07 ms.tex
previously uncommitted changes

\subsection{Derivation of the Steady-State Solution}
Now, consider a sphere of radius $R$ and uniform density, with line-center optical depth $\tau_0 = (k/\sqrt{\pi}\Delta)R$. After changing variables from $\nu$ to $\sigma$ using $d\sigma = \sqrt{2/3}d\nu/(\Delta^2 \phi)$ in Eq. \ref{eq:rteqn} and performing all angular integrals, we obtain Eq. \ref{app:rteqn_derivation}\ref{eq:finaleqn} (see Appendix \ref{app:rteqn_derivation}) and set $p=\alpha_{\rm abs}=0$ to eliminate photon destruction terms. With a photon emission term given by Eq. \ref{app:rteqn_derivation}\ref{eq:jem}, this gives the transfer equation
\be
\nabla^2 J + \left( \frac{k}{\Delta} \right)^2 \frac{\partial^2 J}{\partial \sigma^2} & = & 
- \frac{ \sqrt{6} kL}{4\pi \Delta^2} \delta^3(\vec{x} - \vec{x}_s) \delta (\sigma - \sigma_{\rm s})
\label{eq:rt_no_destr}
\ee
where the form of the source term is in Eq. \ref{app:rteqn_derivation}\ref{eq:jem_v2}. The boundary condition is that there is no incoming flux at the surface, which can be written
\citep{1986rpa..book.....R}
\be
J & = & \sqrt{3} H
\label{eq:bc}
\ee
at $r=R$. An analytic solution $J_d$ which is divergent at the center of the sphere and at the emission frequency may be found. However, it is not a good approximation to the true solution, as it is too large at $r=R$ by a factor of $J_{\rm d}(R,\sigma)/ H_{\rm d}(R,\sigma) \sim a\tau_0/x^2 \sim (a\tau_0)^{1/3} \gg 1$. This solution is included in Fig. \ref{fig:sol_mc_residual_0} for qualitative understanding, but will not be discussed further.

@@

The Monte Carlo method is used to solve the transfer equation numerically as a point of reference for these analytic solutions. This method is valid at arbitrarily large optical depths, being restricted only by its computational demand which grows in proportion to the number of photons used. For each simulation, a total of $2^{20} \approx 10^6$ photon packets are initialized at a monochromatic source frequency $x_s$ and are allowed to propagate through the medium until escaping, at which point their positions, outgoing angles, and frequencies are stored. A constant temperature of $T=10000\ \rm K$ is set for the gas. Frequency redistribution is calculated at each scattering, including the effects of recoil, to obtain the spectrum at the surface of the spherical simulation domain. In the comparisons shown in this section, the raw photon data is binned in escape angle and frequency to obtain spectra. 64 bins are used to ensure good statistics and minimal noise in each bin. Further details of the Monte Carlo simulation are discussed in \cite{2017ApJ...851..150H}.

@@

The size of the frequency-dependent boundary condition correction is dependent on the line center optical depth, which also dictates the x-position of the peak of spectrum according to Eq. \ref{eq:tau_peak_scaling}. This term contributes a significant amount of flux correction even at line center optical depths of $10^7$ where the $J=0$ solution is expected to perform its best, i.e., the majority of photons are pushed further out into the wing and thus the simplifying assumptions made in the solution of the differential equation are better founded. One factor of $(a\tau_0)^{1/3}$ has been scaled out of the x-axis of Fig. \ref{fig:sol_mc_tau} such that the peaks of the distributions are vertically aligned. This scaling has also been applied to the y-axis to preserve normalization of the escape probability. From Eq. \ref{eq:hbc_scaling}, the size of the term $H_{\rm bc}$ should become smaller with larger optical depths, following a $(a\tau_0)^{-1/3}$ scaling. Indeed, agreement between $H_{0 + \rm bc}$ and the Monte Carlo points in Fig. \ref{fig:sol_mc_tau} improves as the optical depth becomes larger. At lower optical depths, the scattering of photons within the Doppler core of the line becomes important, yet our solution is not constructed to model this behavior precisely. The effects of core scattering can be seen in the Monte Carlo data, however, as the peaks of the distribution shift substantially outward in frequency from the line core for $\tau_0 = 10^5$.

@@

Since the solutions have been generalized to allow for a monochromatic source of photons at a frequency $\sigma_s$ away from line center, their agreement with Monte Carlo as a function of source frequency can also be characterized. Photons initialized closer to the line wing may travel larger mean free paths as the opacity is sensitively dependent on the line profile. The larger spatial diffusion contributes to greater escape probability, biasing the frequency distribution of escaping photons. In the limit that $|\rm x_s|$ becomes large, the distribution becomes a delta function in frequency as all photons escape the gas without scattering. Fig. \ref{fig:sol_mc_xinit} shows calculations performed for three initial photon frequencies $\rm x_s = 0.0, 6.0$, and $12.0$ for a line center optical depth of $\tau_0=10^7$. The bias in frequency distribution is slight for $\rm x_s = 6.0$ where the line profile is still relatively large and emitted photons remain trapped in the core, but the increase in spatial diffusion is much more substantial when photons are initialized with $\rm x_s=12.0$, further away from the line core. On this point, we note that the discrepancy between the Monte Carlo data and $H_0$ (effectively the scale of the corrective $H_{\rm bc}$ term) becomes larger as $\rm x_s$ increases. Thus, for sufficiently large optical depths and emission away from line center, the importance of enforcing frequency-dependent boundary conditions at the surface of the sphere grows. The bias in the escaping spectrum is much more noticeable at lower optical depths, shown in Fig. \ref{fig:sol_mc_xinit_lowtau}, where a $\Delta x_s$ of 6 is a significantly larger fraction of the spectral peak location $(a\tau_0)^{1/3}$. Here it is more clear that as the photon initialization frequency becomes further from the optically thick core, the spectrum approaches a delta function in frequency at $\sigma_s$.

\begin{figure}
    \centering
    \includegraphics{xinit.pdf}
    \caption{$H_{\rm bc}$ solutions shown at three line center optical depths, $\tau_0=10^5, 10^7,$ and $10^9$, for photon source frequencies $\sigma_s=0.0, \tau_0, 2\tau_0,$ etc. Thicker lines indicate larger optical depth, and darker lines indicate a photon source frequency further from line center. The axes are scaled to show that the size of the corrective factor agrees with the predicted $(a\tau_0)^{-1/3}$ scaling.}
    \label{fig:xinit}
\end{figure}

Fig. \ref{fig:xinit} shows how the boundary term scales with both optical depth and photon source frequency. The photon initialization frequency is shifted in integer multiples of $\tau_0$. For clarity, only the positive side of the spectrum is shown. From Eq. \ref{eq:hbc_scaling}, the size of the correction should become smaller with larger optical depths, following a $(a\tau_0)^{-1/3}$ scaling. This factor has been scaled out of the figure such that solutions for different optical depths should show close agreement in scale on the figure if the relation holds, even while changing the source frequency of the photons. Indeed, the solutions converge as the optical depth becomes larger. The remaining discrepancy present is due to the width of the Doppler core with respect to the peak of the spectrum; at lower depths, this causes our assumptions about the functional form of the line profile to break down.

@@

We seek now to tabulate all eigenmodes $J_{nm}(\sigma)$ and eigenfrequencies $\gamma_{nm}$ for all spatial $n$ and for all eigenmodes $m$ at that given $n$. A resonant response in $J(n,\sigma,-i\gamma)$ is produced when the imaginary frequency is negative and has no real component. To measure the size of the response to detect where resonances occur, we sum the absolute value of $J(n,\sigma,-i\gamma)$ over $\sigma$ in array form. Call this response $f$, and use the index $j$ to represent the value of the response at discrete points over a range of $\gamma$. In places where $f_j > f_{j-1}$ and $f_j>f_{j+1}$, we have bounded a resonance that occurs in the interval $(\gamma_{j-1},\gamma_{j+1})$. We then call the ``solve'' step to hone in on the eigenvalue before continuing the sweep in $\gamma$. In the solve step, we evaluate the eigenfunctions at the points $(\gamma_{j-1},\gamma_{j},\gamma_{j+1})$ to obtain the responses $f_{j-1}$, $f_j$, and $f_{j+1}$. A ``guess'' at the correct eigenvalue $\gamma_{nm}$ can be calculated from
\be
\gamma_{\rm guess} &=& \frac{a\gamma_{j-1} - \gamma_{j+1}}{a - 1}
\ee
where
\be
a &=& \left(\frac{f_{j} - f_{j-1}}{f_{j} - f_{j+1}}\right)\left(\frac{\gamma_{j+1}-\gamma_{j}}{\gamma_{j-1}-\gamma_{j}}\right).
\ee
This is repeated by replacing former points $(\gamma_{j-1},\gamma_{j},\gamma_{j+1})$ with closer estimates while the size of the response grows as the resonance is approached. Given a sufficiently accurate eigenvalue $\gamma_{nm}$, we now find the corresponding eigenfunction $J_{nm}(\sigma)$. We can write down an approximate expression for the response near the resonances as
\be
J(n,\sigma,-i\gamma) & \simeq & \frac{J_{nm}(\sigma)}{\gamma_{nm}-\gamma} + C(\sigma),
\ee
where $C(\sigma)$ is weakly dependent on $\gamma$, and so we treat it as a constant. We can now solve for $J_{nm}(\sigma)$ by comparing two points near the resonance:
\be
J(n,\sigma,-i\gamma_1)  - J(n,\sigma,-i\gamma_2)& = &  \frac{J_{nm}(\sigma)}{\gamma_{nm}-\gamma_1} -  \frac{J_{nm}(\sigma)}{\gamma_{nm}-\gamma_2}
\ee
where $C(\sigma)$ has cancelled out in the difference. We can immediately solve as
\be
J_{nm}(\sigma) & \simeq & \frac{ J(n,\sigma,\gamma_1)  - J(n,\sigma,\gamma_2) }{ 1/(\gamma_{nm}-\gamma_1) - 1/(\gamma_{nm}-\gamma_2)}.
\ee
\begin{figure}
    \centering
    \includegraphics{Jsoln_n1_m100.pdf}
    \caption{Eigenfunctions $J_{nm}(x)$ for the lowest-order spatial eigenmode $n=1$, and a range of frequency eigenmodes up to $m=100$. Their oscillatory forms have varying amplitudes which sum to create the final form of the mean intensity. The largest contribution always comes from the ($n=1, m=1$) lowest-order eigenfunction. \todo{Show dimensional factor out front of the eigenfunctions to explain why they are so small. Dimensional analysis of the mean intensity itself. Radius, energy (set to 1), delta, and dimensionless parameters a and tau0.}}
    \label{fig:jsoln}
\end{figure}
A set of eigenfunctions $J_{nm}(\sigma)$ are plotted in Fig. \ref{fig:jsoln} to illustrate their relative scales for different frequency eigenmodes $m$ at a fixed spatial eigenmode $n$. Additional terms add smaller, faster-oscillating components that lead to higher accuracy upon summation with the lower-order terms. The oscillations of various modes must cancel perfectly to recreate the form of the solution in Eq. \ref{eq:Jrsigmat}, so many modes $m$ and $n$ are required for sufficient numerical precision in the final result. Too few (or imprecisely-calculated) modes creates ``ringing'' in the function that indicates the presence of imperfect cancellations in the positive and negative contributions of individual poles.

@@

Integrating $dE/dt d\nu$ from Eq. \ref{eq:dEdtdnu} over all times should give a distribution for the emitted frequencies. It is expected that this time-integrated spectrum of the response to an impulse will agree well with the solution for the $J=0$ steady-state spectrum (Eq. \ref{eq:H0surf}). Dividing by the energy $E$, we find the probability distribution
\be \label{eq:spectrum}
P(\nu) & = &  -\frac{16\pi^2 R}{3k\phi E}  \sum_{nm} (-1)^n \gamma_{nm}^{-1} J_{nm}(\sigma).
\ee
Integrating over $\nu$ gives unity if the sum rule in Eq. \ref{eq:sumrule} is true. In Fig. \ref{fig:steadystate}

\begin{figure}
    \centering
    \includegraphics{steadystate.pdf}
    \caption{Comparison of steady-state and time-integrated spectra for 20 spatial eigenmodes and 500 frequency eigenmodes. The accuracy of the red scatter points is limited by the number of frequency eigenmodes summed. The accuracy of the core behavior for the orange solid curve and red scatter points is limited by the number of spatial eigenmodes summed, which is why both solutions fail in the core. \todo{Steady state, Steady state (partial sum), Fluence of time-dependent solution, start at 8 and go to 30. Make the red line thinner and plot it in the background.}}
    \label{fig:my_label}
\end{figure}

@@

The distribution of the times it takes each photon to escape the sphere is calculated from Eq. \ref{eq:dEdtdnu} by integrating over all frequencies. We obtain
\be
P(t)  & = & - \sqrt{\frac{3}{2}} \frac{16\pi^2 R \Delta^2 }{3kE}     \sum_{nm} (-1)^n  e^{s_{nm}t} \int d\sigma J_{nm}(\sigma) 
\nonumber \\ & = &  - \sum_{nm} P_{nm} s_{nm} e^{s_{nm}t},
\label{eq:waittime}
\ee
which is also normalized to unity.

>>>>

\subsection{Derivation of the Steady-State Solution}
Now, consider a sphere of radius $R$ and uniform density, with line-center optical depth $\tau_0 = (k/\sqrt{\pi}\Delta)R$. After changing variables from $\nu$ to $\sigma$ using $d\sigma = \sqrt{2/3}d\nu/(\phi \Delta^2)$ in Eq. \ref{eq:rteqn} and performing all angular integrals, we obtain Eq. \ref{app:rteqn_derivation}\ref{eq:finaleqn} (see Appendix \ref{app:rteqn_derivation}) and set $p=\alpha_{\rm abs}=0$ to eliminate photon destruction terms. With a photon emission term given by Eq. \ref{app:rteqn_derivation}\ref{eq:jem}, this gives the transfer equation
\be
\nabla^2 J + \left( \frac{k}{\Delta} \right)^2 \frac{\partial^2 J}{\partial \sigma^2} & = & 
- \frac{ \sqrt{6} kL}{4\pi \Delta^2} \delta^3(\vec{x} - \vec{x}_s) \delta (\sigma - \sigma_{\rm s})
\label{eq:rt_no_destr}
\ee
where the form of the source term is in Eq. \ref{app:rteqn_derivation}\ref{eq:jem_v2}. The boundary condition is that there is no incoming flux at the surface, which can be written
\citep{1986rpa..book.....R}
\be
J & = & \sqrt{3} H
\label{eq:bc}
\ee
at $r=R$. An analytic solution $J_d$ which is divergent at the center of the sphere and at the emission frequency may be found. However, it is not a good approximation to the true solution, as it is too large at $r=R$ by a factor of $J_{\rm d}(R,\sigma)/ H_{\rm d}(R,\sigma) \sim a\tau_0/x^2 \sim (a\tau_0)^{1/3} \gg 1$. This solution is included in Fig. \ref{fig:sol_mc_residual_0} for qualitative understanding, but will not be discussed further.

@@

The Monte Carlo method is used to solve the transfer equation numerically as a point of reference for these analytic solutions. This method is valid at arbitrarily large optical depths, being restricted only by its computational demand which grows in proportion to the number of photons used. For each simulation, a total of $2^{20} \approx 10^6$ photon packets are initialized at a monochromatic source frequency $x_s$ and are allowed to propagate through the medium until escaping, at which point their positions, outgoing angles, and escape frequencies are stored. A constant temperature of $T=10000\ \rm K$ is set for the gas. Frequency redistribution is calculated at each scattering, including the effects of recoil, to obtain the spectrum at the surface of the spherical simulation domain. In the comparisons shown in this section, the raw photon data is binned in escape angle and frequency to obtain spectra. 64 bins are used to ensure good statistics and minimal noise in each bin. Further details of the Monte Carlo simulation are discussed in \cite{2017ApJ...851..150H}.

@@

The size of the frequency-dependent boundary condition correction is dependent on the line center optical depth, which also dictates the x-position of the peak of spectrum according to Eq. \ref{eq:tau_peak_scaling}. This term contributes a significant amount of flux correction even at line center optical depths of $10^7$ where the $J=0$ solution is expected to perform its best, i.e., the majority of photons are pushed further out into the wing and thus the simplifying assumptions made in the solution of the differential equation are better founded. One factor of $(a\tau_0)^{1/3}$ has been scaled out of the x-axis of Fig. \ref{fig:sol_mc_tau} such that the peaks of the distributions are vertically aligned. This scaling has also been applied to the y-axis to preserve normalization of the escape probability. From Eq. \ref{eq:hbc_scaling}, the size of the term $H_{\rm bc}$ should become smaller with larger optical depths, following a $(a\tau_0)^{-1/3}$ scaling. Indeed, agreement between both analytic flux calculations and the Monte Carlo points in Fig. \ref{fig:sol_mc_tau} improves with $H_{\rm bc}$ providing a fractionally smaller correction to $H_0$ as the optical depth becomes larger. At lower optical depths, the scattering of photons within the Doppler core of the line becomes important, yet our solution is not constructed to model this behavior precisely. The effects of core scattering can be seen in the Monte Carlo data, however, as the peaks of the distribution shift substantially outward in frequency from the line core for $\tau_0 = 10^5$.

@@

Since the solutions have been generalized to allow for a monochromatic source of photons at a frequency $\sigma_s$ away from line center, their agreement with Monte Carlo as a function of source frequency can also be characterized. Photons initialized closer to the line wing may travel larger mean free paths as the opacity is sensitively dependent on the line profile. The larger spatial diffusion contributes to greater escape probability, biasing the frequency distribution of escaping photons. In the limit that $|\rm x_s|$ becomes large, the distribution becomes a delta function in frequency as all photons escape the gas without scattering. Fig. \ref{fig:sol_mc_xinit} shows calculations performed for three initial photon frequencies $\rm x_s = 0.0, 6.0$, and $12.0$ for a line center optical depth of $\tau_0=10^7$. The bias in frequency distribution is slight for $\rm x_s = 6.0$ where the line profile is still relatively large and emitted photons remain trapped in the core, but the increase in spatial diffusion is much more substantial when photons are initialized with $\rm x_s=12.0$, further away from the line core. It is seen here that the discrepancy between the Monte Carlo data and $H_0$ (effectively the scale of the corrective $H_{\rm bc}$ term) becomes larger as $\rm x_s$ increases. Thus, for sufficiently large optical depths and emission away from line center, the importance of enforcing frequency-dependent boundary conditions at the surface of the sphere grows. The bias in the escaping spectrum is much more noticeable at lower optical depths, shown in Fig. \ref{fig:sol_mc_xinit_lowtau}, where a $\Delta x_s$ of 6 is a significantly larger fraction of the spectral peak location $(a\tau_0)^{1/3}$. Here it is more clear that as the photon initialization frequency becomes further from the optically thick core, the spectrum approaches a delta function in frequency at $\sigma_s$.

\begin{figure}
    \centering
    \includegraphics{xinit.pdf}
    \caption{$H_{\rm bc}$ solutions shown at three line center optical depths, $\tau_0=10^5, 10^7,$ and $10^9$, for photon source frequencies $\sigma_s=0.0, \tau_0, 2\tau_0$. Thicker lines indicate larger optical depth, and darker lines indicate a photon source frequency further from line center. The axes are scaled to show that the size of the corrective factor agrees with the predicted $(a\tau_0)^{-1/3}$ scaling.}
    \label{fig:xinit}
\end{figure}

Fig. \ref{fig:xinit} shows how the boundary term scales with both optical depth and photon source frequency. The photon initialization frequency (in units of $\sigma$, as in the differential equation) is shifted in integer multiples of $\tau_0$. For clarity, only the positive side of the spectrum is shown. From Eq. \ref{eq:hbc_scaling}, the size of the correction should become smaller with larger optical depths, following a $(a\tau_0)^{-1/3}$ scaling. This factor has been scaled out of the figure such that solutions for different optical depths and the same source frequency $\sigma_s$ should show close agreement in scale on the figure if the relation holds. Indeed, the scaled solutions converge as the optical depth becomes larger, indicating general agreement with the $(a\tau_0)^{1/3}$ scaling. The remaining discrepancy present is due to the width of the Doppler core with respect to the peak of the spectrum; at lower depths, this causes the assumptions about the functional form of the line profile to break down. From this, we conclude that the errors introduced by the incorrect separation of variables in \cite{1973MNRAS.162...43H}, \cite{1990ApJ...350..216N}, \cite{2006ApJ...649...14D} and others are proportional to $(a\tau_0)^{1/3}$, where $a$ is the ratio of natural to Doppler broadening and $\tau_0$ is the optical depth at line center. An additional amount of error of order $(a\tau_0)^{1/3}$ is introduced as the photon frequency $\sigma_s$ from line center is increased by a factor of $\tau_0$.

@@

We seek now to tabulate all eigenmodes $J_{nm}(\sigma)$ and eigenfrequencies $\gamma_{nm}$ for all spatial $n$ and for all eigenmodes $m$ at that given $n$. A resonant response in $J(n,\sigma,-i\gamma)$ is produced when the imaginary frequency is negative and has no real component. To measure the size of the response to detect where resonances occur, we sum the absolute value of $J(n,\sigma,-i\gamma)$ over $\sigma$ in array form. Call this response $f$, and use the index $j$ to represent the value of the response at discrete points over a range of $\gamma$. In places where $f_j > f_{j-1}$ and $f_j>f_{j+1}$, we have bounded a resonance that occurs in the interval $(\gamma_{j-1},\gamma_{j+1})$. We then call the ``solve'' step to hone in on the eigenvalue before continuing the sweep in $\gamma$. In the solve step, we evaluate the eigenfunctions at the points $(\gamma_{j-1},\gamma_{j},\gamma_{j+1})$ to obtain the responses $f_{j-1}$, $f_j$, and $f_{j+1}$. A ``guess'' at the correct eigenvalue $\gamma_{nm}$ can be calculated from
\be
\gamma_{\rm guess} &=& \frac{b\gamma_{j-1} - \gamma_{j+1}}{b - 1}
\ee
where
\be
b &=& \left(\frac{f_{j} - f_{j-1}}{f_{j} - f_{j+1}}\right)\left(\frac{\gamma_{j+1}-\gamma_{j}}{\gamma_{j-1}-\gamma_{j}}\right).
\ee
This is repeated by replacing former points $(\gamma_{j-1},\gamma_{j},\gamma_{j+1})$ with closer estimates while the size of the response grows as the resonance is approached. Given a sufficiently accurate eigenvalue $\gamma_{nm}$, we now find the corresponding eigenfunction $J_{nm}(\sigma)$. We can write down an approximate expression for the response near the resonances as
\be
J(n,\sigma,-i\gamma) & \simeq & \frac{J_{nm}(\sigma)}{\gamma_{nm}-\gamma} + C(\sigma),
\ee
where $C(\sigma)$ is weakly dependent on $\gamma$, and so we treat it as a constant. We can now solve for $J_{nm}(\sigma)$ by comparing two points near the resonance:
\be
J(n,\sigma,-i\gamma_1)  - J(n,\sigma,-i\gamma_2)& = &  \frac{J_{nm}(\sigma)}{\gamma_{nm}-\gamma_1} -  \frac{J_{nm}(\sigma)}{\gamma_{nm}-\gamma_2}
\ee
where $C(\sigma)$ has cancelled out in the difference. We can immediately solve as
\be
J_{nm}(\sigma) & \simeq & \frac{ J(n,\sigma,\gamma_1)  - J(n,\sigma,\gamma_2) }{ 1/(\gamma_{nm}-\gamma_1) - 1/(\gamma_{nm}-\gamma_2)}.
\ee
\begin{figure}
    \centering
    \includegraphics{Jsoln_n1_m100.pdf}
    \caption{Eigenfunctions $J_{nm}(x)$ for the lowest-order spatial eigenmode $n=1$, and a range of frequency eigenmodes up to $m=100$. Their oscillatory forms have varying amplitudes which sum to create the final form of the mean intensity. The largest contribution always comes from the ($n=1, m=1$) lowest-order eigenfunction.}
    \label{fig:jsoln}
\end{figure}
A set of eigenfunctions $J_{nm}(\sigma)$ are plotted in Fig. \ref{fig:jsoln} to illustrate their relative scales for different frequency eigenmodes $m$ at a fixed spatial eigenmode $n$. The overall scale of the $J_{nm}(\sigma)$ are set by the factor $E/(kR^3)$ with $E$ arbitrarily set to 1, $k = \tau_0 \sqrt{\pi} \Delta / R$, and $R = 10^{11}$ cm. With this normalization, an eigenfunction has typical size ${\sim} E a / \left(R^2 \Delta \right) = 10^{-37}$. Additional terms add smaller, faster-oscillating components that lead to higher accuracy upon summation with the lower-order terms. The oscillations of various modes must cancel perfectly to recreate the form of the solution in Eq. \ref{eq:Jrsigmat}, so many modes $m$ and $n$ are required for sufficient numerical precision in the final result. Too few (or imprecisely-calculated) modes creates ``ringing'' in the function that indicates the presence of imperfect cancellations in the positive and negative contributions of individual poles.

@@

Integrating $dE/dt d\nu$ from Eq. \ref{eq:dEdtdnu} over all times should give a distribution for the emitted frequencies. It is expected that this time-integrated spectrum of the response to an impulse will agree well with the solution for the $J=0$ steady-state spectrum (Eq. \ref{eq:H0surf}). Dividing by the energy $E$, we find the probability distribution
\be \label{eq:spectrum}
P(\nu) & = &  -\frac{16\pi^2 R}{3k\phi E}  \sum_{nm} (-1)^n \gamma_{nm}^{-1} J_{nm}(\sigma).
\ee
Integrating over $\nu$ gives unity if the sum rule in Eq. \ref{eq:sumrule} is true. In Fig. \ref{fig:steadystate}, the resulting spectrum is shown for a sum up to $n=20$ and $m=500$ (labelled ``Time-integrated'') as compared with two analytic solutions: the steady-state $J=0$ solution with no time dependence (``Steady State'') and the expected result for summing \textit{only} spatial modes in the eigenfunction expansion with the time dependence ignored (``Partial Sum''). This graphic illustrates the physics modeled by each variable in the expansion. For example, the spatial modes contribute to the solutions' accuracy in the core of the line. If more spatial modes were calculated, the agreement between the lines extends further into the core. If additional frequency modes are calculated, faster-oscillating terms are incorporated into the sum over the imaginary frequency variable $\gamma$ which create more perfect cancellations with the lower-order terms, reducing the ``ringing'' seen in the time-integrated spectrum. Modeling the physics deep into the line core could have an impact on the accuracy of the escape time distribution, but this would only affect late times since core photons are trapped for a longer duration before escaping.

\begin{figure}
    \centering
    \includegraphics{steadystate.pdf}
    \caption{Comparison of steady-state and time-integrated spectra for 20 spatial eigenmodes and 500 frequency eigenmodes. The accuracy of the time-integrated solution is limited by the number of frequency eigenmodes summed. The accuracy of the core behavior for both the partial sum and the time-integrated solution is limited by the number of spatial eigenmodes summed, which is why both solutions fail in the core.}
    \label{fig:steadystate}
\end{figure}

@@

The distribution of the times it takes each photon to escape the sphere is calculated from Eq. \ref{eq:dEdtdnu} by integrating over all frequencies. We obtain
\be
P(t)  & = & \sqrt{\frac{3}{2}} \frac{16\pi^2 R \Delta^2 }{3kE}     \sum_{nm} (-1)^n  e^{-\gamma_{nm}t} \int d\sigma J_{nm}(\sigma) 
\nonumber \\ & = &  \sum_{nm} P_{nm} \gamma_{nm} e^{-\gamma_{nm}t},
\label{eq:waittime}
\ee
which is also normalized to unity.
